---
title: "PD Preliminary Report"
author: "| Analysts: Kaitlin Olson and Angela Fought \n|\n| Investigators: Samantha
  Holden\n|\n| Report generated: `r format(Sys.Date(), '%m/%d/%Y')`\n"
output:
  html_document:
    highlight: espresso
    number_sections: yes
    theme: yeti
    toc: yes
    toc_depth: 2
    toc_float: yes
  pdf_document:
    toc: yes
    toc_depth: '2'
---
```{r, echo=FALSE, out.width='70%', fig.show='hold'}
knitr::include_graphics("sph_cida_wm_blk (002).png")
```

# Introduction

We received one excel spreadsheet containing information about 46 individuals in a Parkinson's Disease study for their baseline visit and their follow-up visit. Our goal in this project is to explore the relationship between ??? (to come from Samantha)

This report presents the descriptive statistics on the "Year1Analysis" dataset. The resulting dataset had 92 observations, two for each patient.

# Methods

## Data Cleaning

After inspection of the demographic and expected outcome variables, there was very little missing data. Specifically, there was only one missing value for the cognitive composite score outcome.

## Data Analysis

We are waiting on information from Samantha to create hypotheses and an analysis plan.

# Results

## Descriptive Statistics

There were 46 patients that had data recorded for baseline and then follow-up visits (Table 1). At baseline, the median age was 70 (64.0, 74.0) with a range of [53.0,83.0]. There were 16 females and 30 males that participated in this study. Years of education for these individuals ranged from [10,21] with a median of 18 years (16.0,18.0). The median duration of disease experienced by this sample was 3.0 years (2.0,7.8) but this ranged from [0,42]. Descriptive statistics for possible outcome variables at both baseline and follow-up visits are provided in Table 1. Additionally, there were 11 patients that experienced a change in cognitive classification between baseline and follow-up. The changes of classification for these patients can be found in Table 2.

```{r,echo=FALSE}
#read data in
pd <- read.csv("/Users/olsonkai/Desktop/Year1Analysis.csv")
```

# make spaghetti plots

### Angela's number : 734 - 277 - 0926

```{r,echo=FALSE}
#length(unique(pd$redcap_id))
#####demographics
age <- pd$demo_age
sex <- pd$demo_sex #0 female, 1 male
education <- pd$demo_ed
dis.duration <- pd$demo_disdur #disease duration

#####outcomes (totals)
upsa.tot <- pd$upsa_totalscore #combines financial,communication,comprehension,transportation,household management
pdaq.tot <- pd$pdaq_total #difficulty with daily tasks -- reading,time,instructions,discussion,memory
pdcfrs.tot <- pd$pdcfrs_total #more daily tasks -- money,organization,transportation,understanding
updrs.tot <- pd$updrs_total #motor skill questions
moca.tot <- pd$moca_total #executive,naming,attention,language,memory
drs.tot <- pd$drs2_grandtotal #test unsure of what
aes.score <- pd$aes_score #mental health
pdq.tot <- pd$pdq_tot #mobility,emotional,social support,cognition,communication,stigma,bodily discomfort
comp.tot <- pd$comp_tot #cognitive composite score
cog.class <- pd$cogclass_list #cognitive classification
cog.conv <- pd$cogconv #cognitive conversion from previous visit

#####outcomes (other forms)
upsa.z <- pd$upsa_zscore
upsa.base <- pd$upsa_baseline
upsa.change <- pd$upsa_change
pdaq.change <- pd$pdaq_change
pdcfrs.change <- pd$pdcfrs_change
updrs.change <- pd$updrs_change
moca.change <- pd$moca_change
drs.change <- pd$drs2_change
```

### Table 1. Summary statistics for study participants.
```{r, results="asis",echo=FALSE}
library(arsenal)

#####demographics
visit <- pd$visit #0 baseline, 1 followup
age <- pd$demo_age
sex <- pd$demo_sex #0 female, 1 male
education <- pd$demo_ed
dis.duration <- pd$demo_disdur #disease duration

#####outcomes (totals)
upsa.tot <- pd$upsa_totalscore #combines financial,communication,comprehension,transportation,household management
pdaq.tot <- pd$pdaq_total #difficulty with daily tasks -- reading,time,instructions,discussion,memory
pdcfrs.tot <- pd$pdcfrs_total #more daily tasks -- money,organization,transportation,understanding
updrs.tot <- pd$updrs_total #motor skill questions
moca.tot <- pd$moca_total #executive,naming,attention,language,memory
drs.tot <- pd$drs2_grandtotal #test unsure of what
aes.score <- pd$aes_score #mental health
pdq.tot <- pd$pdq_tot #mobility,emotional,social support,cognition,communication,stigma,bodily discomfort
comp.tot <- pd$comp_tot #cognitive composite score
cog.class <- pd$cogclass_list #cognitive classification
cog.conv <- pd$cogconv #cognitive conversion from previous visit
 
#changing factor level labels prior to table construction
sex <- factor(pd$demo_sex,
              labels=c("Female","Male"))
visit <-factor(pd$visit,
               labels=c("Baseline","Follow-Up"))
cog.class <- factor(pd$cogclass_list,
                    labels=c("Normal Cognition","PD-MCI","PDD"))
cog.conv <- factor(pd$cogconv,
                   labels = c("No","Yes"))

#selecting how I want categorical/numeric data displayed #digits is significant figures
mycontrols  <- tableby.control(numeric.stats=c("Nmiss","medianq1q3","range"),
                               cat.stats=c("Nmiss","countpct"),
                               stats.labels=list(Nmiss='N Missing',medianq1q3='Median (Q1,Q3)', range='Range'),
                               digits=1,test = FALSE,total=FALSE)
 
#actual function to create table #demo_sex before ~ splits it by sex
table1 = tableby(visit~ age + sex + education + dis.duration + upsa.tot + pdaq.tot + pdcfrs.tot + updrs.tot + moca.tot + drs.tot + aes.score + pdq.tot + comp.tot, control = mycontrols)
 
#labels for table
mylabels <- list(age = "Age (Years)",sex="Sex",education="Education (Years)",dis.duration="Disease Duration (Years)",upsa.tot="UPSA-2 Total Score",pdaq.tot="PDAQ Total Score",pdcfrs.tot="PD-CFRS Total Score",updrs.tot="UPDRS Part III Total Score",moca.tot="MoCa Total Score",drs.tot="Total DRS Raw Score",aes.score="AES-C Total Score",pdq.tot="Total PDQ-39 Score",comp.tot="Cognitive Composite Score")
 
#displays table with labels and title
summary(table1, labelTranslations = mylabels)
```

### Table 2. Cognitive classification changes that happened from baseline to follow-up.
```{r, echo=FALSE, error=FALSE, warning=FALSE, message=FALSE}
library(dplyr)
library(kableExtra)

#select for patients who had their cognitive classification change from baseline to followup
cog.change <- pd[pd$cogconv==1,c("redcap_id","cogclass_list")]
cog.change <- na.omit(cog.change)
cog.change <- subset(pd, redcap_id %in% c(cog.change$redcap_id))
cog.change$cog.class <- as.factor(cog.change$cogclass_list)
cog.change <- select(cog.change, c(visit,redcap_id,cog.class))
levels(cog.change$cog.class) <- c("Normal Cognition","PD-MCI","PDD")

#make table with 1 row for each patient
base.cog <- cog.change[cog.change$visit==0,]
fu.cog <- cog.change[cog.change$visit==1,]
cog.table <- left_join(base.cog,fu.cog,by="redcap_id")[,c(2,3,5)]

#make table of pre and post cognitive classifications
colnames(cog.table) <- c("Patient ID","Baseline Cognitive Class.","Follow-Up Cognitive Class.")
kable(cog.table) %>%
  kable_styling(bootstrap_options = "striped",full_width = FALSE)
```

<!-- footer -->

---

```{r, echo=FALSE, out.width='70%', fig.show='hold'}
knitr::include_graphics("sph_cida_wm_blk (002).png")
```